pipeComp, a general framework for the evaluation of computational pipelines, reveals performant single-cell RNA-seq preprocessing tools

### 2020-05-04 draft.tex
Updates from Overleaf

Different approaches have been proposed to select the appropriate number of dimensions, from the visual identification of an 'Elbow' (inflexion point) of the variance explained, to more complex algorithms. We evaluated the performance of dimensionality estimators implemented in the \textit{intrinsicDimension} package \cite{johnssonDimensionality2015}, as well as common procedures such as the `elbow` method (inflexion point in the variance explained by each component), some scRNAseq-specific methods such as the JackStraw procedure \cite{ChungJackstraw2015} or \textit{scran}'s \texttt{denoisePCA} \cite{LunScran2016}, and the recent application of Fisher Separability analysis \cite{AlberganteSepar2019}.

@@

All analyses were performed through the \href{https://github.com/plger/pipeComp}{pipeComp} R package, which implements the pipeline framework described here. scVI results and results of the clustering section were produced with \textit{pipeComp} 0.99.4, while all other results were produced with version 0.99.2 (\textit{biorxiv}) branch of the repository). All code to reproduce the simulations and figures is available in the \url{https://github.com/markrobinsonuzh/scRNA\_pipelines\_paper} repository, which also includes the basic datasets with a standardized annotation.

@@

\begin{table}[h!]
\caption{Overview of the benchmark datasets.}
\label{tab:table1}
\begin{tabular}{rlrrl}
  \hline
dataset & source & protocol & description \\ 
  \hline
Koh & GSE85066 & SMARTer & FACS purified H7 hESC in different differention stages \\ 
  Kumar & GSE60749 & SMARTer & Mouse ESC cultured in different conditions \\ 
  Zhengmix4eq & \cite{duoClustering2018} & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix4uneq & \cite{duoClustering2018} & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix8eq & \cite{duoClustering2018} & 10x & Mixtures of FACS purified PBMCs \\ 
  mixology10x3cl & \cite{tianMixology2018} & 10x & Mixture of 3  cancer cell lines from CellBench \\ 
  mixology10x5cl & \cite{tianMixology2018} & 10x & Mixture of 5 cancer cell lines from CellBench \\ 
  simMix1 & - & 10x-based & Simulation of 10 human cell subpopulations \\
  simMix2 & - & 10x-based & Simulation of 9 mouse cell subpopulations \\
   \hline
\end{tabular}
\end{table}

>>>>

Different approaches have been proposed to select the appropriate number of dimensions, from the visual identification of an `elbow' (inflexion point) of the variance explained by each component, to more complex algorithms. We evaluated the performance of dimensionality estimators implemented in the \textit{intrinsicDimension} package \cite{johnssonDimensionality2015}, as well as common procedures such as the elbow method, some scRNAseq-specific methods such as the JackStraw procedure \cite{ChungJackstraw2015} or \textit{scran}'s \texttt{denoisePCA} \cite{LunScran2016}, and the recent application of Fisher Separability analysis \cite{AlberganteSepar2019}.

@@

All analyses were performed through the \href{https://github.com/plger/pipeComp}{pipeComp} Bioconductor package, which implements the pipeline framework described here. scVI results and results of the clustering section were produced with \textit{pipeComp} 0.99.4, while all other results were produced with version 0.99.2 (\textit{biorxiv}) branch of the repository). All code to reproduce the simulations and figures is available in the \url{https://github.com/markrobinsonuzh/scRNA\_pipelines\_paper} repository, which also includes the basic datasets with a standardized annotation.

@@

\begin{table}[h!]
\caption{Overview of the benchmark datasets.}
\label{tab:table1}
\begin{tabular}{rlrrl}
  \hline
dataset & source & # cells & protocol & description \\ 
  \hline
Koh & GSE85066 & 531 & SMARTer & FACS purified differentiation stages \\ 
  Kumar & GSE60749 & 246 & SMARTer & Mouse ESC cultured in different conditions \\ 
  Zhengmix4eq & \cite{duoClustering2018} & 3994 & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix4uneq & \cite{duoClustering2018} & 6498 & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix8eq & \cite{duoClustering2018} & 3994 & 10x & Mixtures of FACS purified PBMCs \\ 
  mixology10x3cl & \cite{tianMixology2018} & 902 & 10x & Mixture of 3  cancer cell lines from CellBench \\ 
  mixology10x5cl & \cite{tianMixology2018} & 3918 & 10x & Mixture of 5 cancer cell lines from CellBench \\ 
  simMix1 & - & 2500 & 10x-based & Simulation of 10 human cell subpopulations \\
  simMix2 & - & 3000 & 10x-based & Simulation of 9 mouse cell subpopulations \\
   \hline
\end{tabular}
\end{table}
